### 2025-03-06 paper-en.tex
Copy paper to new template

% \documentclass[14pt]{mmcs-article}
\documentclass[a4paper,fleqn]{cas-sc}
\usepackage{amsmath, amsthm, amsfonts, amssymb}

\graphicspath{{paper_images/}}
\newtheorem{theorem}{Theorem}
\newtheorem{notice}{Remark}

\newcounter{definition}[section]
\newenvironment{definition}[1][]{\refstepcounter{definition}\par\medskip
    \noindent \textbf{Definition~\thedefinition. #1} \rmfamily
}
{\medskip}

\begin{document}
\shorttitle{Metagraph approach for construction of an (m, n)-regular bipartite graph with a given girth}

\renewcommand{\mod}[1]{\textrm{mod}\ #1}

%%%%% TODO

% Разобрать комментарии по тексту

% Abstract 250 слов

\section{Literature.}

% Список литературы: ~15 штук
% Брать ссылки из цитат
% Все цитаты привести к правильному виду

\begin{thebibliography}{0}

% 1
\bibitem{zemor}
G. Zémor, On Expander Codes, IEEE Trans. on Information theory, IT-47
No 2, (2001) pp. 835–837.

% 2
\bibitem{johnson}
Johnson, Sarah. (2008). Introducing Low-Density Parity-Check Codes. 
% ~-- University of Newcastle, Australia, 2006.

% 3
\bibitem{gallager}
R. Gallager, "Low-density parity-check codes," in IRE Transactions on Information Theory, vol. 8, no. 1, pp. 21-28, January 1962, doi: 10.1109/TIT.1962.1057683

% 4
\bibitem{protographs}
J. Thorpe,
Low-density parity-check (LDPC) codes constructed from protographs.
~-- JPL, IPN Progress Rep., Aug. 2003, vol. 42–154.

% 5
\bibitem{skorohodov_reachability_problem}
Skorokhodov V.A. Generalization of the Reachability Problem on Directed Graphs
~-- Mathematics and Statistics, Vol. 8, No. 6, pp. 699 - 704, 2020. DOI: 10.13189/ms.2020.080610

% 6
\bibitem{tutte_graph_definition}
W.T. Tutte, Graph Theory. Cambridge University Press, Cambridge, 2001, 360 pp.
% https://books.google.rs/books?id=uTGhooU37h4C&printsec=frontcover#v=onepage&q&f=false

% 7
\bibitem{margulis_construction}
Margulis, G. A. (1982). Explicit constructions of graphs without short cycles and low density codes. Combinatorica, 2(1), 71–78. 
doi:10.1007/bf02579283
%% https://sci-hub.se/10.1007/bf02579283

% 9
%% Критика метода Маргулиса
\bibitem{margulis_critique}
David J.C. MacKay, Michael S. Postol, Weaknesses of Margulis and Ramanujan-Margulis Low-Density Parity-Check Codes, Electronic Notes in Theoretical Computer Science, Volume 74, 2003, Pages 97-104, ISSN 1571-0661, doi:10.1016/S1571-0661(04)80768-0.
% https://www.inference.org.uk/mackay/margulis.pdf

% 8
\bibitem{imrich_construction}
Imrich, W. (1984). Explicit construction of regular graphs without small cycles. Combinatorica, 4(1), 53–59. doi:10.1007/bf02579157 
%% https://sci-hub.se/10.1007/bf02579157

% 10
\bibitem{ldpc_encoding}
T. J. Richardson and R. L. Urbanke, "Efficient encoding of low-density parity-check codes," in IEEE Transactions on Information Theory, vol. 47, no. 2, pp. 638-656, Feb 2001, doi: 10.1109/18.910579

% 11
\bibitem{algebraic_aproach_to_sc_ldpc}
Beemer, A., Habib, S., Kelley, C. A., & Kliewer, J. (2017). A generalized algebraic approach to optimizing SC-LDPC codes. In 55th Annual Allerton Conference on Communication, Control, and Computing, Allerton 2017 (pp. 672-679). (55th Annual Allerton Conference on Communication, Control, and Computing, Allerton 2017; Vol. 2018-January). Institute of Electrical and Electronics Engineers Inc. doi:10.1109/ALLERTON.2017.8262802

% https://web.njit.edu/~jkliewer/wp/paper/BHKK_Allerton17.pdf

% 12
\bibitem{lifts_of_graphs}
C. A. Kelley, "On codes designed via algebraic lifts of graphs," 2008 46th Annual Allerton Conference on Communication, Control, and Computing, Monticello, IL, USA, 2008, pp. 1254-1261, doi: 10.1109/ALLERTON.2008.4797704.

% https://typeset.io/pdf/on-codes-designed-via-algebraic-lifts-of-graphs-34h5sq5koc.pdf

% 13
\bibitem{vasic_combinatorial_construction}
B. Vasic and O. Milenkovic, "Combinatorial constructions of low-density parity-check codes for iterative decoding," in IEEE Transactions on Information Theory, vol. 50, no. 6, pp. 1156-1176, June 2004, doi: 10.1109/TIT.2004.828066.
% https://repository.arizona.edu/bitstream/handle/10150/641965/Combinatorial%20Constructions%20of%20Low-DensityParity-Check%20Codes%20for%20Iterative%20Decoding.pdf;jsessionid=9F016719B821C7D7CEBB9C48D7B37CB1?sequence=1

%14
\bibitem{Kim2002ExplicitCO}
Kim, Jon-Lark, Uri N. Peled, Irina G. Perepelitsa and Vera Pless. “Explicit Construction of Families of LDPC Codes with Girth at Least Six.” (2002).

%https://homepages.math.uic.edu/~pless/preprint/allerton.pdf

%% https://errorcorrectionzoo.org/

%% Discard

% 13

% Слишком общо, но можно по ссылкам пройтись
% \bibitem{quantum_error_correction}
% G. Zémor, “On Cayley Graphs, Surface Codes, and the Limits of Homological Coding for Quantum Error Correction”
% https://www.researchgate.net/publication/220776025_On_Cayley_Graphs_Surface_Codes_and_the_Limits_of_Homological_Coding_for_Quantum_Error_Correction

% \bibitem{metagraphs_and_applications}
% A. Basu, R.W. Blanning. Mategraphs and Their Applications
% ~-- %https://books.google.rs/books?id=9DlhUIJklqUC&printsec=frontcover#v=onepage&q&f=false

\end{thebibliography}

\end{document}